---
title: "Fetch GSOD Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fetch and Import GSOD Weather Data

Using the `get_GSOD()` function from _GSODR_, download and reformat
Global Surface Summary of the Day (GSOD) weather data from the US National
Centers for Environmental Information (NCEI) for the years 1980 to 2016 between
latitudes -60 and 60 only by using the `agroclimatology` argument in
`get_GSOD()`. This process will take several hours depending on Internet
connection and processor speed to download and clean the data. The resulting
comma separated values (CSV) files vary in size from a few hundred megabytes to
more than half a gigabyte.

```{R get_GSOD, eval=TRUE, echo=TRUE, message=FALSE, results='hide'}
get_write <- function(year_list, dsn) {
  weather <- GSODR::get_GSOD(years = year_list,
                             max_missing = 5,
                             agroclimatology = TRUE)

  fname <- paste0("GSOD_", weather[1, "YEAR"], ".csv")

  readr::write_csv(weather, path = file.path(dsn, fname), na = "NA")
  
  # clean up and free up RAM/swap as this never completed in the first attempt
  rm(weather)
  gc()
}

# generate a vector of "years" from 1983 to 2016
year_list <- as.list(seq(from = 1983, to = 2016, by = 1))

# use purrr's map() to apply the get_write() function to the list of years
# download (get) and create (write) CSV files of the GSOD weather data
lapply(X = year_list, FUN = get_write, dsn = "~/Data/GSOD")

```

******

# Appendices

## R Session Information

```{r system information, echo=FALSE}

devtools::session_info()

```
