---
title: "Fetch GSOD Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fetch and Import GSOD Weather Data

Using the `get_GSOD()` function from _GSODR_, download and reformat
Global Surface Summary of the Day (GSOD) weather data from the US National
Centers for Environmental Information (NCEI) for the years 1980 to 2016 between
latitudes -60 and 60 only by using the `agroclimatology` argument in
`get_GSOD()`. This process will take several hours depending on Internet
connection and processor speed to download and clean the data. Data are then
subset to include only the fields of interest, a station id value, longitude,
latitude, elevation, date, temperature and relative humidity data. The resulting
comma separated values (CSV) files vary in size from a few hundred megabytes to
more than half a gigabyte. However, by saving them as .bz2 files, it saves
several gigabytes of storage space. Note that `fname` below ends in a .bz2 file
extension so `readr::write_csv` automatically compresses the files.

```{R get_GSOD, eval=TRUE, echo=TRUE, message=FALSE, results='hide'}
fetch_GSOD <- function(year_list, dsn) {

  # fetch GSOD data from NCEI server
  weather <- GSODR::get_GSOD(years = year_list,
                             max_missing = 5,
                             agroclimatology = TRUE)

  # select only the fields that are necessary for or to be interpolated
  # this saves more than 1/2 the space of the full original data in storage
  weather <-
    weather[, c(
      "STNID",
      "LON",
      "LAT",
      "ELEV_M_SRTM_90m",
      "YEAR",
      "YDAY",
      "TEMP",
      "MAX",
      "MIN",
      "RH")]

  # remove any stations lacking elevation data,
  # they cannot be used in calculations
  weather <- weather[!is.na(weather$ELEV_M_SRTM_90m),]

  # create YEAR object for naming file out
  YEAR <- weather$YEAR[1]

  # create file name
  fname <- paste0("GSOD_", YEAR, ".bz2")

  # write a compressed CSV file to disk in the specified location
  readr::write_csv(weather, path = file.path(dsn, fname), na = "NA")

  # clean up and free up RAM/swap
  rm(weather)
  gc()
}

# generate a vector of "years" from 1983 to 2016
year_list <- as.list(seq(from = 1983, to = 2017, by = 1))

# lapply to apply the get_write() function to the list of years
# download (get) and create (write) CSV files of the GSOD weather data
lapply(X = year_list, FUN = fetch_GSOD, dsn = "~/Data/GSOD")
```

******

# Appendices

## R Session Information

```{r system information, echo=FALSE}

devtools::session_info()

```
