---
title: "Fetch GSOD Data"
author: "Adam H Sparks - Centre for Crop Health, USQ"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

  knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
  )
```

# Fetch, Import and Process Elevation Data

Download Shuttle Radar Topography Mission (SRTM) digital elevation model (DEM)
data from the Worldclim Database [(link)](http://www.worldclim.org/) to use in
the interpolation as an elevation covariate. Set data type to INT2S since the
.bil file is unsigned, set -9999 to NA, then aggregate to 1 arc-degree to match
NASA - POWER. Since the DEM covers more area than the agroclimatology data from
the _GSODR_ package, crop the DEM to match the extent before plotting.

```{r Alt, echo=TRUE, fig.cap="Elevation plot of 1deg SRTM data", fig.fullwidth = TRUE, dev.args=list(pointsize=6), eval=TRUE}
# load libraries
library(raster)
library(ggplot2)
library(viridis)
library(rgdal)

# set up workspace
tf.zip <- tempfile()

download.file(
  "http://biogeo.ucdavis.edu/data/climate/worldclim/1_4/grid/cur/alt_10m_bil.zip",
  destfile = tf.zip,
  mode = "wb"
)
unzip(tf.zip, exdir = tempdir()) # unzip downloaded file
z <-
  raster(paste0(tempdir(), "/alt.bil")) # import elevation raster object
dataType(z) <- "INT2S"

# aggregate the SRTM data
z <- aggregate(z, 6) # aggregate up to 1deg to match NASA - POWER resolution
z[z == -9999] <- NA # set -9999 to NA

z # inspect the object

# crop SRTM data at -60/60 for agroclimatology only
z <- crop(z, 
          c(
            xmin = -180,
            xmax = 180,
            ymin = -60,
            ymax = 60
          )
)

z_spdf <- as(z, "SpatialPixelsDataFrame")
z_df <- as.data.frame(z_spdf)
colnames(z_df) <- c("value", "x", "y")
```

## Plot the DEM

```{r, eval = TRUE}
ggplot() +
  geom_tile(data = z_df, aes(x = x, y = y, fill = value)) +
  scale_fill_viridis() +
  xlab("LON") +
  ylab("LAT") +
  theme_bw() +
  coord_map()
```

### Save the Spatial File to Disk

Using the _rgdal_ GeoPackage driver, we will save the file to disk for use in
interpolating the data, using the spatial data frame object.

```{r save_file, eval=TRUE}
writeRaster(
  z,
  file = "~/Data/SRTM",
  overwrite = TRUE
)
```

******

# Appendices

## R Session Information

```{r system information, echo=FALSE}

devtools::session_info()

```
