---
title: "GRID"
author: "Adam H Sparks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GRID}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(GRID)
```

## Using fetch_DEM

Fetch a digital elevation model (DEM) from WorldClim data, aggregate to 1 arc
degree and crop at -60/60 degrees latitude and return a spatial object in the
current R session.

```{r R_DEM, eval=FALSE}
DEM <- fetch_DEM()
```

Fetch a digital elevation model (DEM) from WorldClim data and aggregate to 1 arc
degree.

```{r dsn_DEM, eval=FALSE}
DEM <- fetch_DEM()
```

Fetch a digital elevation model (DEM) from WorldClim data, aggregate to 15 arc
minutes and crop at -60/60 degrees latitude and save a GeoTIFF file of the
resulting DEM to local disk.

```{r dsn_DEM_0.25, eval=FALSE}
DEM <- fetch_DEM(resolution = .25, dsn = "~/GSOD")
```

## Using fetch_GSOD

Fetch GSOD data for 1998 to 2000.

```{r fetch_GSOD, eval=FALSE}
gsod <- fetch_GSOD(years = 1998:2000)
```

## Using interpolate_gsod with a list of data frames


## Using interpolate_gsod with a list of CSV files

# Example 1: Saving all Files to Local Disk

```{r local_disk, eval=FALSE}

```

# Example 2: Using In-Memory Data Only

```{r in_memory, eval=FALSE}
# Create a list of GSOD files
gsod_list <- fetch_GSOD(years = 1998:1999)

# Fetch and aggregate the raster digital elevation model
dem <- fetch_DEM()

# Run the function for MAX and MIN temperature
lapply(X = gsod_list, FUN = interp, dem = dem, vars = c("MAX", "MIN"))
```
