---
title: "GSOD Interpolation as an Improved Alternative To NASA POWER Agroclimatology Data"
author: "Adam H. Sparks - Centre for Crop Health, USQ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Global Surface Summary of the Day - GSOD data,
[(link)](https://data.noaa.gov/dataset/global-surface-summary-of-the-day-gsod)
is free, ground-based (or bouy) weather station data with global
coverage. When properly cleaned and interpolated this data can provide a better
alternative to NASA/POWER for agroclimatology work. This document describes how
to reproduce these data using freely available data, software and methodologies.

# Install and Load the Necessary R Packages
```{r InstallPackages, echo=TRUE, message=FALSE, eval=FALSE}
library(GSODR)
library(doParallel)
library(foreach)
library(ggplot2)
library(raster)
library(rastervis)
library(readr)
library(rgdal)
```

# Fetch and Import Weather Data

Using the `get_GSOD()` function from _GSODR_, download and reformat
Global Surface Summary of the Day (GSOD) weather data from the US National
Climatic Data Center (NCDC) for the year 2010 between latitudes -60 and 60
only. This process will take several minutes or hours depending on Internet
connection and processor speed to download and clean the data. The resulting
comma separated values (CSV) file is roughly 465mb.

```{r get_GSOD, eval=TRUE, echo=TRUE, message=FALSE}
GSOD <- get_GSOD(years = 2010:2011, max_missing = 5, agroclimatology = TRUE)
```

Inspect the data frame of the GSOD data and plot the station locations included in the data.

```{r import GSOD, eval=TRUE, echo=TRUE, message=FALSE}
summary(GSOD)

```

```{r plot_stations, eval=TRUE, echo=TRUE, fig.cap="Locations of station data in dataset", fig.fullwidth = TRUE, dev.args=list(pointsize=6)}
# plot station locations
ggplot(data = GSOD, aes(y = LAT, x = LON)) +
  geom_point() +
  xlab("LON") +
  ylab("LAT") +
  theme_bw() +
  coord_map()
```

# Fetch, Import and Process Elevation Data

Download Shuttle Radar Topography Mission (SRTM) digital elevation model (DEM)
data from the Worldclim Database [(link)](http://www.worldclim.org/) to use in
the interpolation as an elevation covariate. Set data type to INT2S since the
.bil file is unsigned, set -9999 to NA, then aggregate to 1 arc-degree to match
NASA/POWER. Since the DEM covers more area than the agroclimatology data from
the `GSODR` package, crop the DEM to match the extent before plotting.

```{r Alt, echo=TRUE, fig.cap="Elevation plot of 1deg SRTM data", fig.fullwidth = TRUE, dev.args=list(pointsize=6), cache=TRUE}
# set up workspace
tf.zip <- tempfile()

download.file(
  "http://biogeo.ucdavis.edu/data/climate/worldclim/1_4/grid/cur/alt_10m_bil.zip",
  destfile = tf.zip, mode = "wb")
unzip(tf.zip, exdir = tempdir()) # unzip downloaded file
z <- raster(paste0(tempdir(), "/alt.bil")) # import elevation raster object
dataType(z) <- "INT2S"

# aggregate the SRTM data
z <- aggregate(z, 6) # aggregate up to 1deg to match NASA/POWER resolution
z[z == -9999] <- NA # set -9999 to NA

z # inspect the object

# crop SRTM data at -60/60 for agroclimatology only
agro <- extent(c(xmin = -180, xmax = 180, ymin = -60, ymax = 60))
z <- crop(z, agro)

z_spdf <- as(z, "SpatialPixelsDataFrame")
z_df <- as.data.frame(z_spdf)
colnames(z_df) <- c("value", "x", "y")

ggplot() +
  geom_tile(data = z_df, aes(x = x, y = y, fill = value)) +
  scale_fill_viridis() +
  xlab("LON") +
  ylab("LAT") +
  theme_bw() +
  coord_map()

```

# Daily Weather Data Cleaning, Thin Plate Splining and Interpolation

GSOD temperature data are checked for consistency using `boxplot.stats()` to
identify and remove daily outliers for both TMAX and TMIN, separately.

Once the GSOD data are cleaned, thin plate splining, the `Tps()` function, from
 _fields_ is used to create an object that can then be used to create
an interpolated surface of temperatures using the SRTM DEM with the
`interpolate()` function from _raster_.

Use _foreach_ to iterate through the data paralellising the process.

1. check for outliers in GSOD data and remove them,

2. interpolate creating a global surface from the values,

3. write GeoTIFF files to disk for every day, every weather variable specified.

This is an extremly processor and time-intensive process for the entire global
dataset. It is suggested to use a computer dedicated just to this task, while it
will run on a fairly modest desktop, the computer may become unresponsive while
performing this operation.

```{r tps_interpolation, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, cache=FALSE}
cl <- makeCluster(detectCores() - 2)
registerDoParallel(cl)
YEARMODA <- 20100101 # set this to c(2010:2011) to run all days, all years. Only one day here as example
itx <- iter(YEARMODA)
ity <- iter(c("TEMP", "MIN", "MAX"))
i <- NULL
j <- NULL

foreach(i = itx, .packages = c("fields", "raster", "foreach")) %dopar% {
  foreach(t = ity) %do% {
    j <- GSOD[GSOD$YEARMODA == i, ][, c(8:9, 11, 19, 33, 35)]
    j <- j[complete.cases(j), ]
    
    # create object with x, y and weather var
    gsod_temp <- na.omit(data.frame(j$LON, j$LAT, j$ELEV_M_SRTM_90m,
                                    as.numeric(paste0(j[[t]]))))
    
    # remove outliers
    temp_bxs <- boxplot.stats(gsod_temp[, 4])
    temp <- gsod_temp[!gsod_temp[, 4] %in% temp_bxs$out, ]
    
    # create correction dataset
    temp_vals <- temp[, 4]
    names(temp_vals) <- NULL
    temp_xyz <- temp[, 1:3]
    
    # create thin plate spline object
    tps_temp <- fields::Tps(temp_xyz[, c("j.LON", "j.LAT", "j.ELEV_M_SRTM_90m")],
                            temp_vals, lon.lat = TRUE)
    
    # interpolate thin plate spline object and write to disk
    tps_pred_temp <- interpolate(z, tps_temp, xyOnly = FALSE)
    
    writeRaster(tps_pred_temp,
                filename = paste0("~/GTiff/GSOD_SRTM.90m_", t, "_", i, ".tiff"),
                format = "GTiff", dataType = "INT2S",
                options = c("COMPRESS=LZW", "TFW=YES"),
                overwrite = TRUE)
  }
}

stopCluster(cl)

```
******

# Appendices

## Computer Information  

> **iMac (21.5-inch, Late 2015)**  
> **Processor** 2.8 GHz Intel Core i5  
> **Memory** 8 GB 1867 MHz DDR3  

## R Session Information

```{r system information, echo=FALSE}

sessionInfo()

```
