---
title: "glint Use Case"
author: "Adam H. Sparks - Centre for Crop Health, USQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GRID Use Case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(glint)
```

# Introduction

Global Surface Summary of the Day - GSOD data,
[(link)](https://data.noaa.gov/dataset/global-surface-summary-of-the-day-gsod)
is free, ground-based (or buoy) weather station data with global
coverage. When properly cleaned and interpolated this data can provide a better
alternative to NASA/POWER for agroclimatology work. This document describes how
to reproduce these data using freely available data, software and methodologies.

# Daily Weather Data Cleaning, Thin Plate Splining and Interpolation

The _glint_ package makes this process as simple as possible. Described below are
the steps that the package uses internally to create gridded weather data from
the GSOD data.

i## Downloading GSOD Data

_glint_ offers customised functionality to download GSOD data, `get_GSOD()`
using functionality from the [_GSODR_](https://ropensci.github.io/GSODR/)
package. _glint_ will provide only the data necessary for interpolation,
excluding the other data that GSOD provides while omitting any station that has
more than a maximum five missing days to help ensure data quality. The four
variables that _glint_ can interpolate are:

* TEMP - Mean daily temperature,

* MAX - Daily maximum temperature,

* MIN - Daily minimum temperature,

* RH - Mean daily relative humidity.

#k# Downloading the Digital Elevation Model

_glint_ offers customised functionality, `make_DEM()` to download a digital
elevation model based on the Shuttle Radar Topography Mission (SRTM) from
[WorldClim](https://wwww.worldclim.org/) and aggregated it to:

* 1 degree (matching historical [NASA - POWER](https://power.larc.nasa.gov/)
data, which could provide rainfall),

* 0.5 degree (matching [NASA - POWER](https://power.larc.nasa.gov/) data,
which can provide rainfall),

* 0.25 degree (matching [CHIRPS](http://chg.geog.ucsb.edu/data/chirps/) data,
which can provide rainfall)

## Interpolating the GSOD Data

_glint_ offers a simple function, `interpolate_GSOD()`, to automate error
checking, interpolation and saving GeoTIFF files of the resulting data.

When using the `interpolate_GSOD()` function, GSOD data are first checked for
consistency using `boxplot.stats()` to identify and remove daily outliers for
any variable requested.

Once the GSOD data are cleaned, thin plate splining, the `Tps()` function, from
 _fields_ is used to create an object that can then be used to create
an interpolated surface of temperatures using the SRTM DEM with the
`interpolate()` function from _raster_.

The data are then returned in a list of raster stacks by weather variable in the
R session or saved to disk in GeoTIFF file format in an INT2S format with LZW
compression for saving disk space.

This is an extremely processor and time-intensive process for the entire global
data set. It is suggested to use a computer dedicated just to this task, while it
will run on a fairly modest desktop, the computer may become unresponsive while
performing this operation.
k
## Example Creating a Set of Global GeoTIFF Files from 1983 to 2010

### Step 1: Download the GSOD data

```{r get_GSOD, eval=FALSE, message='hide', results='hide'}
future::plan("multisession")
years <- as.list(seq(from = 1983, to = 1984, by = 1))
lapply(X = years, FUN = make_GSOD_set, dsn = "/mnt/GSOD")
```

### Step 2: Download the Digital Elevation Model and Aggregate to 0.5 Degree

```{r make_DEM, eval=FALSE, message='hide', results='hide'}
DEM <- make_DEM(dsn = "/mnt/DEM")
```

### Step 3: Interpolate the GSOD Data from On-disk Files

```{r list_GSOD_files, eval=FALSE, message='hide', results='hide'}
file_list <- list.files("/mnt/GSOD", full.names = TRUE)
```

Since `future::plan()` was already set to `multisession` previously, that will
apply here and the interpolation will automatically be done in parallel.

```{r interpolate, eval=FALSE, message='hide', results='hide'}
GRID <- lapply(X = file_list, FUN = interpolate_GSOD, dem = DEM,
               dsn = "/mnt/Cache/GTiff", vars = c("MAX", "MIN"))
```

******

# Appendices

## Reproducibility

```{r system information, echo=FALSE}

sessioninfo::session_info()

```
