---
title: "GRID Use Case"
author: "Adam H. Sparks - Centre for Crop Health, USQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GRID Use Case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Global Surface Summary of the Day - GSOD data,
[(link)](https://data.noaa.gov/dataset/global-surface-summary-of-the-day-gsod)
is free, ground-based (or buoy) weather station data with global
coverage. When properly cleaned and interpolated this data can provide a better
alternative to NASA/POWER for agroclimatology work. This document describes how
to reproduce these data using freely available data, software and methodologies.

# Daily Weather Data Cleaning, Thin Plate Splining and Interpolation

The _GRID_ package makes this process as simple as possible. Described below are
the steps that the package uses internally to create gridded weather data from
the GSOD data.

## Downloading GSOD Data

_GRID_ offers customised functionality to download GSOD data, `get_GSOD()`
using functionality from the [_GSODR_](https://ropensci.github.io/GSODR/)
package. _GRID_ will provide only the data necessary for interpolation,
excluding the other data that GSOD provides while omitting any station that has
more than a maximum five missing days to help ensure data quality. The four
variables that _GRID_ can interpolate are:

* TEMP - Mean daily temperature,

* MAX - Daily maximum temperature,

* MIN - Daily minimum temperature,

* RH - Mean daily relative humidity.

## Downloading the Digital Elevation Model

_GRID_ offers customised functionality, `get_DEM()` to download a digital
elevation model based on the Shuttle Radar Topography Mission (SRTM) from
[WorldClim](https://wwww.worldclim.org/) and aggregated it to:

* 1 degree (matching [NASA - POWER](https://power.larc.nasa.gov/) data, which
can provide rainfall),

* 0.5 degree (matching future [NASA - POWER](https://power.larc.nasa.gov/) data,
which can provide rainfall),

* 0.25 degree (matching [CHIRPS](http://chg.geog.ucsb.edu/data/chirps/) data,
which can provide rainfall)

## Interpolating the GSOD Data

_GRID_ offers a simple function, `interpolate_GSOD()`, to automate error
checking, interpolation and saving GeoTIFF files of the resulting data.

When using the `interpolate_GSOD()` function, GSOD data are first checked for
consistency using `boxplot.stats()` to identify and remove daily outliers for
any variable requested.

Once the GSOD data are cleaned, thin plate splining, the `Tps()` function, from
 _fields_ is used to create an object that can then be used to create
an interpolated surface of temperatures using the SRTM DEM with the
`interpolate()` function from _raster_.

The data are then returned in a list of raster stacks by weather variable in the
R session or saved to disk in GeoTIFF file format in an INT2S format with LZW
compression for saving disk space.

This is an extremely processor and time-intensive process for the entire global
data set. It is suggested to use a computer dedicated just to this task, while it
will run on a fairly modest desktop, the computer may become unresponsive while
performing this operation.

## Example Creating a Set of Global GeoTIFF Files from 1983 to 2016

### Step 1: Download the GSOD data

```{r get_GSOD, eval=FALSE, message='hide', results='hide'}
years <- as.list(seq(from = 1983, to = 2016, by = 1))
lapply(X = years, FUN = get_GSOD, dsn = "~/Data/GSOD")
```

### Step 2: Download the Digital Elevation Model and Aggregate to 1 Degree

```{r get_DEM, eval=FALSE, message='hide', results='hide'}
DEM <- get_DEM()
```
### Step 3: Interpolate the GSOD Data from On-disk Files

```{r list_GSOD_files, eval=FALSE, message='hide', results='hide'}
file_list <- list.files("~/Data/GSOD", full.names = TRUE)
```

For a Unix-type operating system, it is possible to run the process in parallel
to reduce the time necessary to generate all variables. This example shows how
to run the function and save GTiff files in "~/Cache/GTiff" using two cores.

```{r interpolate_2_cores,  eval=FALSE, message='hide', results='hide'}
GRID <- lapply(X = file_list, FUN = interpolate_GSOD, dem = dem,
dsn = "~/Cache/GTiff", vars = c("MAX", "MIN"), cores = 2)
```

#### Using a Single Core on Windows

If a Windows operating system is detected, the function will automatically
default to using a single core for the processes as Windows does not support
the parallel procedures that it uses. No need to specify the cores here, will
default to one. If set to higher than one, it will automatically reset to one to
run.

```{r interpolate_windows,  eval=FALSE, message='hide', results='hide'}
GRID <- lapply(X = file_list, FUN = interpolate_GSOD, dem = dem,
dsn = "~/Cache/GTiff", vars = c("MAX", "MIN"))
```

******

# Appendices

## Reproducibility

```{r system information, echo=FALSE}

sessioninfo::session_info()

```
